Implementation Design Patters for Pursuit-Evasion Problems with Bounded Speed

### 2012-12-01 paper.tex
fixed some typos

\section{Paradigm 1 - Discrete Formulation}
The Discrete Formulation of Pursuit-Evasion studies focus on recognizing scenarios in a set of states on a connected graph. This creates a paradigm where strategies are much more easily recognized and explained. Routes and movements are generalized and win/lose situations for pursuers and evaders can often be predicted ahead of time by the nature of the graph played on.

@@

Let the environment that the pursuers and evaders coexist in be a finite, connected graph \(G\). Let \(G\)'s vertices be represented as \(v_1,v_2,...,v_n\), each being a distinct location in the environment. All edges are considered to be undirected and of equal length. Let the set of all \(q\) pursuers in \(G\) can be represented as \(P = {p_1,...,p_q}\), with each being on some \(v_i\) in \(G\). Subsequently, also let the set of all \(w\) evaders in \(G\) can be represented as \(E = {e_1,...,e_w}\), with each being on some \(v_i\) in \(G\). Any vertex in \(G\) can be occupied by more than one pursuer and/or evader.

@@

There can be varieties on each of the rules for this scenario. The Pursuer(s) and/or evader(s) may not need to necessarily move each turn, which can sometimes determine whether or not one side wins or loses. Also the pursuer(s) or evader(s) may not have perfect information when making decisions, only able to "see" others at certain distances. If pursuers do not know any information other than the node they occupy, we can often represent these scenarios as search games. \cite{parsons}

@@

To start, we will begin by defining the scenario environment as an open set \(R\subset\mathbb{R}^2\). \(R\) is to be a simply-connected and has some sort of polygonal boundary. Some scenarios define that \(R\) has holes. \(P\subset R\) is assumed to be the set of points that represent the locations of the various pursuers of the scenario. Accordingly \(E\subset R\) is assumed to make up the set of evaders, their locations being represented as points as well. Members of \(P\) has perfect information about their location with respect to \(R\) and each other, but not necessarily \(E\). Both pursuers and evaders move at various speeds within \(R\). Members of \(P\) generally are expected to move at a bounded consistent speed \(V_P\), while evaders can be viewed as having either a bounded speed or in some cases, unbounded. \cite{limVis, robotics} Pursuers are generally considered to have "caught" the evaders when being within a certain distance of them.

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{"visPoly"}
\caption{A pursuer P in an example R. The visibility polygon for P is shaded in showing the area P can view. Note the edge gaps that bound the visibility polygon and the rest of R. If P were to move as far to the right as possible, the visible area near to the top of R would vanish.}
\end{figure}

As stated previously, the set of pursuers \(P\) do not have perfect information of \(R\) and can only see a combined limited region at one time. Subsequently, they may not know where the evaders \(E\) are located. This visible region is known as the \emph{visibility polygon} \(V\). \(V\) is composed of a collection of line segments that may partially align with \(R\)'s shape itself. It's assumed \(V\)'s topology changes changes as it's corresponding \(P\) moves about \(R\). The line segments of \(V\) are that cross across the face of \(R\)'s polygonal shape are known as \emph{gap edges}. These gap edges allow us to divide \(R\) into shrinking and growing regions of visibility. \cite{limVis, robotics}

@@

% mathematical definitions and/or data structures
To start, we will begin by defining the scenario environment as an open set \(R\subset\mathbb{R}^2\). \(R\) is to be a simply-connected and has some sort of polygonal boundary. Let sets \(P,E\subset{R}\) be the positions of the pursuers and evaders and their velocities identical to that found in the Continuous Formulation. Let \(G\) be an undirected, weighted-edge graph representing \(R\), with each of its vertex in a state being either \emph{visible, uncontaminated,} or \emph{contaminated}. Later in this section we will explain how \(G\) represents \(R\). Each pursuer and evader will keep track of which vertex it currently "occupies". It is also assumed that all pursuers in \(P\) know their location relative to \(R\).

@@

% delunay triangulation and minimum spanning tree connection
To have \(G\) represent \(R\), we take advantage of \(R\)'s shape being a closed polygon. By performing, a Delaunay Triangulation on \(R\), we acquire a triangulation composed of \(R\)'s vertices. \cite{dTri} It is assumed that the triangles outside of \(R\) are "stripped" and discarded immediately. We are left with a set of adjacent triangles that make up \(R\). For each triangle, add a corresponding vertex to \(G\) and add an edge between each adjacent triangle. The weight of an edge between two triangle's vertices is the distance between each triangle's midpoint in barycentric coordinates. \cite{bCenter} \(G\) now represents a set of general areas in \(R\) and the approximate distances between each.

@@

Contaminated triangles and vertices in \(G\) are meant to represent regions in \(R\) where an evader could possibly be. In that case, if a contaminated region is adjacent to an uncontaminated region, it is possible that an evader could move from the contaminated to the uncontaminated region, making both uncontaminated. This creates a degree of "spreading" contamination that needs to be kept in check by the pursuers. Since the evaders move at a bounded speed though, it can be expected that contamination does not happen necessarily instantaneously. For this reason the edges in \(G\) have a weight associated with them. Contamination from one vertex/triangle to another happens after a delay \(D\) where: \begin{displaymath}
D = \frac{\text{length of edge between vertices}}{\text{maximum bounded speed of evader(s)}}\times \text{S} ; S\in(0,1]
\end{displaymath}\(S\) represents a "shortcut" value, which can be altered to scale contamination speed/ Evaders may not necessarily go from center to center in each triangle, and could spend less time in one than expected.

>>>>

\section{Paradigm 1 - Discrete Formulation}
The Discrete Formulation of Pursuit-Evasion studies focuses on recognizing scenarios in a set of states on a connected graph. This creates a paradigm where strategies are much more easily recognized and explained. Routes and movements are generalized and win/lose situations for pursuers and evaders can often be predicted ahead of time by the nature of the graph played on.

@@

Let the environment that the pursuers and evaders coexist in be a finite, connected graph \(G\). Let \(G\)'s vertices be represented as \(v_1,v_2,...,v_n\), each being a distinct location in the environment. All edges are considered to be undirected and of equal length. Let the set of all \(q\) pursuers in \(G\) be represented as \(P = {p_1,...,p_q}\), with each being on some \(v_i\) in \(G\). Also let the set of all \(w\) evaders in \(G\) be represented as \(E = {e_1,...,e_w}\), with each being on some \(v_i\) in \(G\). Any vertex in \(G\) can be occupied by more than one pursuer and/or evader.

@@

There can be variations on each of the rules for this scenario. The Pursuer(s) and/or evader(s) may not need to necessarily move each turn, which can sometimes determine whether or not one side wins or loses. Also the pursuer(s) or evader(s) may not have perfect information when making decisions, only able to "see" others at certain distances. If pursuers do not know any information other than the node they occupy, we can often represent these scenarios as search games. \cite{parsons}

@@

To start, we will begin by defining the scenario environment as an open set \(R\subset\mathbb{R}^2\). \(R\) is simply-connected and has some sort of polygonal boundary. Some scenarios define that \(R\) has holes. \(P\subset R\) is assumed to be the set of points that represent the locations of the various pursuers of the scenario. Accordingly \(E\subset R\) is assumed to make up the set of evaders, their locations being represented as points as well. Members of \(P\) have perfect information about their location with respect to \(R\) and each other, but not necessarily \(E\). Both pursuers and evaders move at various speeds within \(R\). Members of \(P\) generally are expected to move at a bounded consistent speed \(V_P\), while evaders can be viewed as having either a bounded speed or in some cases, unbounded. \cite{limVis, robotics} Pursuers are generally considered to have "caught" the evaders when being within a certain distance of them.

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{"visPoly"}
\caption{A pursuer P in an example R. The visibility polygon for P is shaded, showing the area P can view. Note the edge gaps that bound the visibility polygon and the rest of R. If P were to move as far to the right as possible, the visible area near to the top of R would vanish.}
\end{figure}

As stated previously, the set of pursuers \(P\) do not have perfect information of \(R\)'s contents and can only see a combined limited region at one time. Subsequently, they may not know where the evaders \(E\) are located. This visible region is known as the \emph{visibility polygon} \(V\). \(V\) is composed of a collection of line segments that may partially align with \(R\)'s shape itself. It's assumed \(V\)'s topology changes changes as it's corresponding \(P\) moves about \(R\). The line segments of \(V\) that cross across the face of \(R\)'s polygonal shape are known as \emph{gap edges}. These gap edges allow us to divide \(R\) into shrinking and growing regions of visibility. \cite{limVis, robotics}

@@

% mathematical definitions and/or data structures
To start, we will begin by defining the scenario environment as an open set \(R\subset\mathbb{R}^2\). \(R\) is to be simply-connected and has some sort of polygonal boundary. Let sets \(P,E\subset{R}\) be the positions of the pursuers and evaders and their velocities identical to those found in the Continuous Formulation. Let \(G\) be an undirected, weighted-edge graph representing \(R\), with each of its vertex in a state being either \emph{visible, uncontaminated,} or \emph{contaminated}. Later in this section we will explain how \(G\) represents \(R\). Each pursuer and evader will keep track of which vertex it currently "occupies". It is also assumed that all pursuers in \(P\) know their location relative to \(R\).

@@

% delunay triangulation and minimum spanning tree connection
To have \(G\) represent \(R\), we take advantage of \(R\)'s shape being a closed polygon. By performing a Delaunay Triangulation on \(R\), we acquire a triangulation composed of \(R\)'s vertices. \cite{dTri} It is assumed that the triangles outside of \(R\) are "stripped" and discarded immediately. We are left with a set of adjacent triangles that make up \(R\). For each triangle, add a corresponding vertex to \(G\) and add an edge between each adjacent triangle. The weight of an edge between two triangle's vertices is the distance between each triangle's midpoint in barycentric coordinates. \cite{bCenter} \(G\) now represents a set of general areas in \(R\) and the approximate distances between each.

@@

Contaminated triangles and vertices in \(G\) are meant to represent regions in \(R\) where an evader could possibly be. In that case, if a contaminated region is adjacent to an uncontaminated region, it is possible that an evader could move from the contaminated to the uncontaminated region, making both contaminated. This creates a degree of "spreading" contamination that needs to be kept in check by the pursuers. Since the evaders move at a bounded speed though, it can be expected that contamination does not happen necessarily instantaneously. For this reason the edges in \(G\) have a weight associated with them. Contamination from one vertex/triangle to another happens after a delay \(D\) where: \begin{displaymath}
D = \frac{\text{length of edge between vertices}}{\text{maximum bounded speed of evader(s)}}\times \text{S} ; S\in(0,1]
\end{displaymath}\(S\) represents a "shortcut" value, which can be altered to scale contamination speed/ Evaders may not necessarily go from center to center in each triangle, and could spend less time in one than expected.
